---
title: "Resumen opendatascot package (in Spanglish)"
author: "Inma"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
options(scipen = 999) #to avoid scientific notations in plot
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# opendatascot

>The opendatascot R package allows you to download directly into R any dataset from the Open access Scotland's official statistics website:  https://statistics.gov.scot/home

First download the package from https://scotgovanalysis.github.io/opendatascot/ writing the code below in the console. (quitándole los hastags, lo escribes en la consola por que sólo se hace una vez para la instalación). 
```{r}
#install.packages("devtools")
#devtools::install_github("jsphdms/opendatascot")
```

```{r libraries}
# Importing libraries
library(opendatascot)
library(tidyverse)
library(knitr)
library(kableExtra)
```

# Lo que me pareció más útil

## See all datasets available:

You can see all datasets available in  https://statistics.gov.scot with the All_datasets() function. For each dataset it shows the URI, the name of the dataset and the publisher. 
```{r}
All_datasets <- ods_all_datasets()
```

## Getting the dataset you want

Retrieving one of the 297 datasets using its URI. (creas un nuevo objeto y ya la tienes en tu enviroment para trabajar con ella como quieras) 
```{r}
low_qual <- ods_dataset("adults-16-64-years-with-low-or-no-qualifications")
head(low_qual) %>%
  kable()
```

## Tidyverse

Now, you can Work with the tidyverse verbs and extract whatever info you need 

```{r}
low_qual %>%
  select(refPeriod, gender, age, value) %>%
  filter(gender != "all" & age !="16-64" & refPeriod == 2019) %>%
  group_by(gender, age) %>%
  summarize(total = sum(value)) %>%
  ggplot(aes(x=age, y=total, fill=gender))+ geom_col() + 
  ggtitle("Adults with low or no qualifications in Scotland in 2019") +
  xlab("age group") + ylab("Total number") 
```



# Otras funciones 

El package tiene más funciones, pero no les veo mucha utilidad porque todo se puede hacer con el tidyverse una vez que te la bajas a R. 
La funcion ods_structure() me pareció que confundía porque en algunas datasets mostraba las categorias distintas a como se ven en la data. 
También hay otras para selecionar la geografía. Te las pongo aqui de todas maneras:

to see the structure of one dataset ods_structure("uri_name")
```{r}
#ods_structure("uri_name")
ods_structure("homelessness-applications")
```

filtering ods_dataset("uri_name", column1="what_you_want", column2=c("what_you_want1, "what_you_want2"))
```{r subseting}
ods_dataset("homelessness-applications",
            applicationType ="all-applications",
            refPeriod = c("2015/2016","2016/2017"))
```

Option for geography are:
"dz" - returns datazones only
"iz" - returns intermediate zones only
"hb" - returns healthboards only
"la" - returns local authorities only
"sc" - returns Scotland as a whole only
```{r geography}
ods_dataset("homelessness-applications",
            geography = "la")

```

ods_find_lower_geographies() will return a dataframe of all geographies that are contained by the geography you pass it 
```{r}
lower <- ods_find_lower_geographies("S02000003")
lower
```

ods_find_higher_geographies() will return a dataframe of all geographies that contain the geography you pass it
```{r}
higher <- ods_find_higher_geographies("S02000003")
higher
```

This dataframes can then be passed to ods_dataset to get information about these geographies! We just need to select the vector of geography codes, and use the refArea filter option:
```{r}
ods_dataset("house-sales-prices",
             refArea = lower$geography,
             measureType = "mean",
             refPeriod = "2013")
```


```{r}
ods_dataset("house-sales-prices",
             refArea = higher$geography,
             measureType = "mean",
             refPeriod = "2013") %>%
  kable()
```

### Este es el problema que encontré con la funcion de ver la estructura:

```{r}
ods_structure("gender-employment-gap")
```
If I use the value of  refArea as it is shown in the previous output; the result is a tibble of 0 x 5

```{r}
ods_dataset("gender-employment-gap",
            refPeriod = "2020-Q1",
            refArea = "Midlothian") 
```

but it works when entering the code for Midlothian (that it isn´t shown in the structure output)
```{r}
ods_dataset("gender-employment-gap",
            refPeriod = "2020-Q1",
            refArea = "S12000019") %>%
  kable()
```
             
             
Also, category age is shown without the  **-** before **years** :  "16-24 years"  but in the data set the format of the value is:   "25-49-years" 

>my github: https://github.com/InmaculadaRM


references: 
https://github.com/DataScienceScotland/opendatascot



